---
title: "Chapter 02: Covariation in data"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter02.R](R/chapter02.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Covariation in data`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Calculation of correlation in R

### Box 01

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-01}
    df <- read.csv("apartment_price_data.csv")
```

### Box 02

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-02}
    xbar <- mean(df$living_area)
    ybar <- mean(df$price)
```

### Box 03

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-03}
    is.na(df$price)
```

### Box 04

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-04}
    !is.na(df$price) & !is.na(df$living_area)
```

### Box 05

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-05}
    df <- df[!is.na(df$price) & !is.na(df$living_area), ]
```

### Box 06

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-06}
    xbar <- mean(df$living_area)
    ybar <- mean(df$price)
    n <- nrow(df)
```

### Box 07

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-07}
      s_xy <- 1 / (n - 1) * sum((df$living_area - xbar) *
                                (df$price - ybar))
```

### Box 08

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-08}
      s_x2 <- 1 / (n - 1) * sum((df$living_area - xbar)^2)
      s_y2 <- 1 / (n - 1) * sum((df$price - ybar)^2)
```

### Box 09

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-09}
      corr_xy <- s_xy / (sqrt(s_x2) * sqrt(s_y2))
```

### Box 10

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-10}
      corr_xy <- cov(df$living_area, df$price) / 
                 (sd(df$living_area) * sd(df$price))
```

### Box 11

Textbook context: Section: Pearson's correlation.

```{r calculation-of-correlation-in-r-11}
      corr_xy <- cor(df$living_area, df$price)
```

## Calculation of Spearman's correlation in R

### Box 12

Textbook context: Section: Spearman's rank correlation coefficient.

```{r calculation-of-spearman-s-correlation-in-r-12}
    df$Rx <- rank(df$living_area)
    df$Ry <- rank(df$price)
```

### Box 13

Textbook context: Section: Spearman's rank correlation coefficient.

```{r calculation-of-spearman-s-correlation-in-r-13}
    mean_Rx <- (n + 1) / 2
    mean_Ry <- (n + 1) / 2
```

### Box 14

Textbook context: Section: Spearman's rank correlation coefficient.

```{r calculation-of-spearman-s-correlation-in-r-14}
    covRxRy <- sum((df$Rx - mean_Rx)*(df$Ry-mean_Ry)) / (n - 1)
   sRx <- sqrt(sum((df$Rx - mean_Rx)^2) / (n - 1))
sRy <- sqrt(sum((df$Ry - mean_Ry)^2) / (n - 1))
```

### Box 15

Textbook context: Section: Spearman's rank correlation coefficient.

```{r calculation-of-spearman-s-correlation-in-r-15}
    rS <- covRxRy / (sRx * sRy)
```

### Box 16

Textbook context: Section: Spearman's rank correlation coefficient.

```{r calculation-of-spearman-s-correlation-in-r-16}
    rS_alt <- cor(df$Rx, df$Ry, method = "spearman")
```

## Regression in R

### Box 17

Textbook context: Section: Simple linear regression | Subsection: Coefficient of determination.

```{r regression-in-r-17}
  beta_1_hat <- s_xy / s_x2
  beta_0_hat <- ybar - xbar * beta_1_hat
```

### Box 18

Textbook context: Section: Simple linear regression | Subsection: Coefficient of determination.

```{r regression-in-r-18}
  y_hat <- beta_0_hat + beta_1_hat * df$living_area
```

### Box 19

Textbook context: Section: Simple linear regression | Subsection: Coefficient of determination.

```{r regression-in-r-19}
 u_hat <- df$price - y_hat
```

### Box 20

Textbook context: Section: Simple linear regression | Subsection: Coefficient of determination.

```{r regression-in-r-20}
  RSS <- sum(u_hat^2)
  TSS <- sum((df$price - ybar)^2)
  R2 <- 1 - RSS / TSS
```

### Box 21

Textbook context: Section: Simple linear regression | Subsection: Coefficient of determination.

```{r regression-in-r-21}
 ols_model <- lm(price ~ living_area, data=df)
```

### Box 22

Textbook context: Section: Simple linear regression | Subsection: Coefficient of determination.

```{r regression-in-r-22}
 summary(ols_model)
```
