---
title: "Chapter 08: Regression analysis with dependent error terms"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter08.R](R/chapter08.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Regression analysis with dependent error terms`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Project Star and clustering in R

### Box 01

Textbook context: Section: Clustered data | Subsection: OLS with clustered data.

```{r project-star-and-clustering-in-r-01}
# Load data
df <- read.csv(file.path("data", "star.csv"))

# NOTE (updated data schema):
# class_type  : class assignment (e.g. "SMALL", "REGULAR", "AIDE")
# class_id    : class/teacher identifier (cluster id)
# read_score  : grade 3 reading score (outcome)
# school_id   : school identifier (optional for multilevel models)

# Create treatment indicator: small_class
# 1 if SMALL, 0 if REGULAR/AIDE, NA otherwise
df$small_class <- ifelse(df$class_type == "SMALL", 1, NA)
df$small_class <- ifelse(df$class_type %in% c("AIDE", "REGULAR"), 0, df$small_class)

# Keep complete cases for outcome, regressor, and cluster id
df <- df[!is.na(df$class_id) & !is.na(df$read_score) & !is.na(df$small_class), ]

# Baseline regression
ols_model <- lm(read_score ~ small_class, data=df)

# Standard deviation of outcome (for standardized effect if desired)
sd(df$read_score, na.rm=TRUE)

# Robust and cluster-robust inference
library(lmtest)
library(sandwich)

# HC0 robust standard errors
coeftest(ols_model, vcov = vcovHC(ols_model, type="HC0"))

# Cluster-robust standard errors (clustered by class_id)
V_cl <- vcovCL(ols_model, cluster = df$class_id, type="HC0")
coeftest(ols_model, vcov = V_cl)

# Birth month example
ols_model2 <- lm(read_score ~ birth_month, data=df)
coeftest(ols_model2, vcov = vcovHC(ols_model2, type="HC0"))

V_cl2 <- vcovCL(ols_model2, cluster = df$class_id, type="HC0")
coeftest(ols_model2, vcov = V_cl2)

# Combined model
ols_model3 <- lm(read_score ~ small_class + birth_month, data=df)
V_cl3 <- vcovCL(ols_model3, cluster = df$class_id, type="HC0")
coeftest(ols_model3, vcov = V_cl3)
```

## Project Star and multilevel models in R

### Box 02

Textbook context: Section: Clustered data | Subsection: Multilevel-models.

```{r project-star-and-multilevel-models-in-r-02}
    install.packages(c("lme4"))
```

### Box 03

Textbook context: Section: Clustered data | Subsection: Multilevel-models.

```{r project-star-and-multilevel-models-in-r-03}
df <- read.csv(file.path("data", "star.csv"))

# Create treatment indicator
df$small_class <- ifelse(df$class_type == "SMALL", 1, NA)
df$small_class <- ifelse(df$class_type %in% c("AIDE", "REGULAR"), 0, df$small_class)

# Keep complete cases
df <- df[!is.na(df$class_id) & !is.na(df$read_score) & !is.na(df$small_class), ]

library(lme4)

# Random intercept by class (class_id)
multi_model1 <- lmer(read_score ~ small_class + (1 | class_id), data = df)
summary(multi_model1)

# Random intercept by class and school (if school_id exists / is non-missing)
# (If school_id has missing values, consider dropping them before fitting.)
df2 <- df[!is.na(df$school_id), ]
multi_model2 <- lmer(read_score ~ small_class + (1 | class_id) + (1 | school_id), data = df2)
summary(multi_model2)
```

## The within estimator with municipal data in R

### Box 04

Textbook context: Section: Panel data | Subsection: Regression with individual fixed effects.

```{r the-within-estimator-with-municipal-data-in-r-04}
df <- read.csv(file.path("data", "municip_data.csv"))
df <- df[df$municip_name != "Gotland", ]

library(plm)
library(lmtest)
library(sandwich)

within_model <- plm(
  tax_rate ~ left_coalition_last_term,
  effect = "individual",
  index = c("municip_name", "year"),
  data = df
)

coeftest(within_model, vcov = vcovHC(within_model, type="HC0", cluster="group"))

# Two-way FE (municipality + year)
twoway_model <- plm(
  tax_rate ~ left_coalition_last_term,
  effect = "twoways",
  index = c("municip_name", "year"),
  data = df
)
coeftest(twoway_model, vcov = vcovHC(twoway_model, type="HC0", cluster="group"))
```
