---
title: "Chapter 06: Multiple Linear Regression"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter06.R](R/chapter06.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Multiple Linear Regression`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Multiple regression in R

### Box 01

Textbook context: Section: The general multiple linear regression model.

```{r multiple-regression-in-r-01}
 df <- read.csv("apartment_price_data.csv")
```

### Box 02

Textbook context: Section: The general multiple linear regression model.

```{r multiple-regression-in-r-02}
 ols_model <- lm(price ~ living_area + monthly_fee +
 new_production, data=df)
```

### Box 03

Textbook context: Section: The general multiple linear regression model.

```{r multiple-regression-in-r-03}
    library("lmtest")
    library("sandwich")
    rob_inf <- coeftest(ols_model,
      vcov = vcovHC(ols_model, type = "HC0"))
```

### Box 04

Textbook context: Section: The general multiple linear regression model.

```{r multiple-regression-in-r-04}
      lmtest::coefci(ols_model,
  vcov. = vcovHC(ols_model, type = "HC0"),
  level = 0.95)
```

### Box 05

Textbook context: Section: The general multiple linear regression model.

```{r multiple-regression-in-r-05}
 b1 <- rob_inf["living_area", "Estimate"]
 se1 <- rob_inf["living_area", "Std. Error"]
 t_stat <- (b1 - 0.04) / se1
 df_model <- ols_model$df.residual
 p_value <- (1 - pt(abs(t_stat), df_model)) * 2
```

## $F$-test in R

### Box 06

Textbook context: Section: $F$-test.

```{r f-test-in-r-06}
 ols_model_l <- lm(price ~ living_area + monthly_fee +
  new_production + build_year, data=df)
 summary(ols_model_l)
```

### Box 07

Textbook context: Section: $F$-test.

```{r f-test-in-r-07}
    n <- nobs(ols_model_l)
```

### Box 08

Textbook context: Section: $F$-test.

```{r f-test-in-r-08}
    b <- !is.na(df$new_production) & !is.na(df$build_year)
```

### Box 09

Textbook context: Section: $F$-test.

```{r f-test-in-r-09}
    ols_model_s <- lm(price ~ living_area + monthly_fee,
                              data=df[b, ])
```

### Box 10

Textbook context: Section: $F$-test.

```{r f-test-in-r-10}
    residuals_l <- residuals(ols_model_l)
    residuals_s <- residuals(ols_model_s)
    RSS_l <- sum(residuals_l^2)
    RSS_s <- sum(residuals_s^2)
```

### Box 11

Textbook context: Section: $F$-test.

```{r f-test-in-r-11}
    K <- 4
    G <- 2
```

### Box 12

Textbook context: Section: $F$-test.

```{r f-test-in-r-12}
    F_obs <- ((RSS_s - RSS_l) / (K - G)) / 
             (RSS_l / (n - (K + 1)))
```

### Box 13

Textbook context: Section: $F$-test.

```{r f-test-in-r-13}
 p <- 1 - pf(F_obs, df1=K-G, df2=n-(K+1))
```

### Box 14

Textbook context: Section: $F$-test.

```{r f-test-in-r-14}
    anova(ols_model_s, ols_model_l)
```

### Box 15

Textbook context: Section: $F$-test.

```{r f-test-in-r-15}
 summary(ols_model_l)
```

## Robust $F$-test in R

### Box 16

Textbook context: Section: $F$-test | Subsection: Robust $F$-test.

```{r robust-f-test-in-r-16}
    library("lmtest")
    library("sandwich")
```

### Box 17

Textbook context: Section: $F$-test | Subsection: Robust $F$-test.

```{r robust-f-test-in-r-17}
    waldtest(ols_model_s, ols_model_l,
             vcov = vcovHC(ols_model_l, type = "HC0"))
```

## Confidence intervals for expected value and prediction intervals in R

### Box 18

Textbook context: Section: Uncertainty of the conditional expectation | Subsection: Prediction interval.

```{r confidence-intervals-for-expected-value-and-prediction-intervals-in-r-18}
 ols_model <- lm(price ~ living_area, data=df)
 x0 <-80
 yhat_x80 <- predict(ols_model,
 newdata=data.frame(living_area=x0))
```

### Box 19

Textbook context: Section: Uncertainty of the conditional expectation | Subsection: Prediction interval.

```{r confidence-intervals-for-expected-value-and-prediction-intervals-in-r-19}
  n <- nobs(ols_model)
  RSS <- sum(residuals(ols_model)^2)
  s2_epsilon <- RSS / (n - 2)
  xbar <- mean(df$living_area)
  SSX <- sum((df$living_area - xbar)^2)
  var_eY_x0 <- s2_epsilon * (1 / n + (x0 - xbar)^2 / SSX)
```

### Box 20

Textbook context: Section: Uncertainty of the conditional expectation | Subsection: Prediction interval.

```{r confidence-intervals-for-expected-value-and-prediction-intervals-in-r-20}
 alpha <- 0.05
 t_crit <- qt(1-alpha/2, df=n-2)
 lb_ci_eY_x0 <- yhat_x80 - t_crit * sqrt(var_eY_x0)
 ub_ci_eY_x0 <- yhat_x80 + t_crit * sqrt(var_eY_x0)
```

### Box 21

Textbook context: Section: Uncertainty of the conditional expectation | Subsection: Prediction interval.

```{r confidence-intervals-for-expected-value-and-prediction-intervals-in-r-21}
 var_tildeepsilon_x0 <- var_eY_x0 + s2_epsilon
 lb_pi_x0 <- yhat_x80 - t_crit * sqrt(var_tildeepsilon_x0)
 ub_pi_x0 <- yhat_x80 + t_crit * sqrt(var_tildeepsilon_x0)
```

### Box 22

Textbook context: Section: Uncertainty of the conditional expectation | Subsection: Prediction interval.

```{r confidence-intervals-for-expected-value-and-prediction-intervals-in-r-22}
 ols_model2 <- lm(price ~ living_area + monthly_fee +
 new_production, data=df)
```

### Box 23

Textbook context: Section: Uncertainty of the conditional expectation | Subsection: Prediction interval.

```{r confidence-intervals-for-expected-value-and-prediction-intervals-in-r-23}
 eval_df <- data.frame(living_area=80,
 monthly_fee=3000, new_production=1)
 predict(ols_model2, newdata = eval_df,
 interval = "confidence",
 level=.95)
 predict(ols_model2, newdata = eval_df,
 interval = "prediction",
 level=.95)
```
