---
title: "Chapter 11: Time series analysis"
subtitle: "Runnable workbook for Regression with R and Python"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This runnable workbook supplements the reference workbook. It adds setup code and lightweight fixes so the chapter can usually be rendered and run end-to-end.

- Runnable companion script: [R/chapter11.R](R/chapter11.R)
- Reference workbook: [chapter11_workbook.Rmd](chapter11_workbook.Rmd)
- Textbook chapter: `Time series analysis`
- Package installation lines are shown as comments and are not run automatically during rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
knitr::opts_knit$set(root.dir = chapter_dir)
```

## Time series analysis with R

### Box 01

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-01}
    df <- read.csv(file.path("data", "time_series_sweden.csv"))
```

### Box 02

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-02}
    gdp_ts <- ts(df$gdp, frequency = 4, start = c(1981, 1))
```

### Box 03

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-03}
    plot(gdp_ts)
```

### Box 04

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-04}
    gdpL1 <- lag(gdp_ts, k=-1)
    gdpL2 <- lag(gdp_ts, k=-2)
    gdpF1 <- lag(gdp_ts, k=1)
    gdpF2 <- lag(gdp_ts, k=2)
```

### Box 05

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-05}
    ma <- (gdpL2/2 + gdpL1 + gdp_ts +  gdpF1 + gdpF2/2)/4
```

### Box 06

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-06}
    plot(ma)
```

### Box 07

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-07}
    mean_gdp <- mean(gdp_ts)
    r_1 <- (sum((gdp_ts - mean_gdp) * (gdpL1 - mean_gdp))) / 
           sum((gdp_ts - mean_gdp)^2)
```

### Box 08

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-08}
 autocorrelation_function <- acf(gdp_ts)
```

### Box 09

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-09}
    plot(autocorrelation_function)
```

### Box 10

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-10}
    dloggdp <- (log(gdp_ts) - log(gdpL1)) * 100
    maL1 <- lag(ma, k=-1)
    dlogma <- (log(ma) - log(maL1)) * 100
```

### Box 11

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-11}
plot(dloggdp)
plot(dlogma)
```

### Box 12

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-12}
 dloggdpL1 <- lag(dloggdp, k=-1)
 dlogmaL1 <- lag(dlogma, k=-1)
```

### Box 13

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-13}
    mv_ts <- ts.union(gdp_ts, gdpL1, gdpL2, gdpF1, gdpF2, 
                      ma, dloggdp, dlogma, dloggdpL1,
                      dlogmaL1)
    dft <- data.frame(mv_ts)
```

### Box 14

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-14}
 ar1_model <- lm(dloggdp ~ dloggdpL1, data=dft)
 summary(ar1_model)
```

### Box 15

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-15}
 library (sandwich)
 cov_matrix <- NeweyWest(ar1_model, lag=5, prewhite=FALSE)
```

### Box 16

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-16}
    std_errors <- sqrt(diag(cov_matrix))
```

### Box 17

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-17}
 ar1_model_alt <- ar.ols(dloggdp,
 order.max = 1,
 aic=FALSE,
 demean = FALSE,
 intercept = TRUE)
```

### Box 18

Textbook context: Section: Autoregressive regression | Subsection: Deterministic or stochastic trend? | Subsubsection: Stochastic Trend.

```{r runnable-time-series-analysis-with-r-18}
    ar5_model <- ar.ols(dloggdp, 
      order.max = 5,
      aic=TRUE,
      demean = FALSE, 
      intercept = TRUE)
```
