---
title: "Chapter 07: Nonlinear functional form"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter07.R](R/chapter07.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Nonlinear functional form`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Air pollution and weather in R

### Box 01

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-01}
    df <- read.csv("pollution_sf.csv")
```

### Box 02

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-02}
 df$wind_cat <- cut(df$wind_direction,
 breaks=seq(0, 360, by=10),
 include.lowest=TRUE)
```

### Box 03

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-03}
    df_agg <- aggregate(df$pm25,
                        by=list(wind_dir_interval=df$wind_cat),
                        FUN=mean)
    df_agg$wind_direction <- seq(5, 355, 10)
```

### Box 04

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-04}
    names(df_agg)[names(df_agg) == "x"] <- "pm25_mean"
```

### Box 05

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-05}
    plot(df_agg$wind_direction, df_agg$pm25_mean)
```

### Box 06

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-06}
    df$north_wind <- ifelse(df$wind_direction > 315 |
                            df$wind_direction <= 45, 1, 0)
    df$east_wind <- ifelse(df$wind_direction > 45 &
                            df$wind_direction <= 135, 1, 0)
    df$south_wind <- ifelse(df$wind_direction > 135 &
                            df$wind_direction <= 225, 1, 0)
    df$west_wind <- ifelse(df$wind_direction > 225 &
                            df$wind_direction <= 315, 1, 0)
    df$land_wind <- ifelse(df$wind_direction > 330 |
                           df$wind_direction <= 150, 1, 0)
    df$strong_wind <- ifelse(df$wind_speed >= 3, 1, 0)
```

### Box 07

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-07}
    df$wind_direction2 <- df$wind_direction^2
    df$wind_direction3 <- df$wind_direction^3
    df$wind_direction4 <- df$wind_direction^4
```

### Box 08

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-08}
    ols_model_poly <- lm(pm25 ~ wind_direction +
      wind_direction2 + wind_direction3 +
      wind_direction4, data=df)
```

### Box 09

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-09}
    ols_model_poly_alt <- lm(pm25 ~ 
      poly(wind_direction, 4), data=df)
```

### Box 10

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-10}
    newdata <- data.frame(wind_direction=seq(0, 360, 1))
```

### Box 11

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-11}
 newdata$pred <- predict(ols_model_poly_alt, newdata)
```

### Box 12

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-12}
    newdata$pred_alt <- coef(ols_model_poly)[1] +
      coef(ols_model_poly)[2] * newdata$wind_direction +
      coef(ols_model_poly)[3] * newdata$wind_direction^2 +
      coef(ols_model_poly)[4] * newdata$wind_direction^3 +
      coef(ols_model_poly)[5] * newdata$wind_direction^4
```

### Box 13

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-13}
 plot(df_agg$wind_direction, df_agg$pm25_mean)
 lines(newdata$wind_direction, newdata$pred)
```

### Box 14

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-14}
 df$land_wind_strong_wind <- df$land_wind * df$strong_wind
 df$land_wind_wind_speed <- df$land_wind * df$wind_speed
```

### Box 15

Textbook context: Section: Example: Relationship between weather and air pollution.

```{r air-pollution-and-weather-in-r-15}
    library("lmtest")
    library("sandwich")
    ols_model1 <- lm(pm25 ~ land_wind, data=df)
    ols_model2 <- lm(pm25 ~ land_wind + strong_wind +
                     land_wind_strong_wind, data=df)
    ols_model3 <- lm(pm25 ~ land_wind + wind_speed +
                     land_wind_wind_speed, data=df)
    coeftest(ols_model1, vcov=vcovHC(ols_model1, type="HC0"))
    coeftest(ols_model2, vcov=vcovHC(ols_model2, type="HC0"))
    coeftest(ols_model3, vcov=vcovHC(ols_model3, type="HC0"))
```
