---
title: "Chapter 05: The simple linear regression model"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter05.R](R/chapter05.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `The simple linear regression model`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Inference for regression in R

### Box 01

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-01}
 s2_epsilon <- RSS / (n - 2)
 SSX <- sum((df$living_area - mean(df$living_area))^2)
 var_hat_beta_1_hat <- s2_epsilon / SSX
```

### Box 02

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-02}
 beta_1_H0 <- 0
 se_beta_1_hat <- sqrt(var_hat_beta_1_hat)
 t_stat <- (beta_1_hat - beta_1_H0) / se_beta_1_hat
 p_val <- 2 * (1 - pt(abs(t_stat), df=n-2))
```

### Box 03

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-03}
    alpha <- 0.1
    t_crit <- qt(1-alpha/2, df=n-2)
    lb <- beta_1_hat - t_crit * se_beta_1_hat
    ub <- beta_1_hat + t_crit * se_beta_1_hat
```

### Box 04

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-04}
 sum_ols_model <- summary(ols_model)
```

### Box 05

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-05}
 reg_table <- sum_ols_model$coefficients
```

### Box 06

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-06}
 reg_table[2, 3]
```

### Box 07

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-07}
 confint(ols_model, level=0.9)
```

### Box 08

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-08}
 ols_model2 <- lm(price ~ new_production, data=df)
```

### Box 09

Textbook context: Section: Inference to the population with spherical error terms.

```{r inference-for-regression-in-r-09}
    confint(ols_model2, level=0.9)
```

## Robust inference in R

### Box 10

Textbook context: Section: Asymptotic inference to the population.

```{r robust-inference-in-r-10}
 dx2 <- (df$living_area - mean(df$living_area))^2
```

### Box 11

Textbook context: Section: Asymptotic inference to the population.

```{r robust-inference-in-r-11}
 var_beta_1_var_r <- sum(dx2 * u_hat^2) / (sum(dx2))^2
```

### Box 12

Textbook context: Section: Asymptotic inference to the population.

```{r robust-inference-in-r-12}
 se_beta_1_hat_r <- sqrt(var_beta_1_var_r)
```

### Box 13

Textbook context: Section: Asymptotic inference to the population.

```{r robust-inference-in-r-13}
    library("lmtest")
    library("sandwich")
```

### Box 14

Textbook context: Section: Asymptotic inference to the population.

```{r robust-inference-in-r-14}
 coeftest(ols_model, vcov=vcovHC(ols_model, type="HC0"))
```
