---
title: "Chapter 05: The simple linear regression model"
subtitle: "Runnable workbook for Regression with R and Python"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This runnable workbook supplements the reference workbook. It adds setup code and lightweight fixes so the chapter can usually be rendered and run end-to-end.

- Runnable companion script: [R/chapter05.R](R/chapter05.R)
- Reference workbook: [chapter05_workbook.Rmd](chapter05_workbook.Rmd)
- Textbook chapter: `The simple linear regression model`
- Package installation lines are shown as comments and are not run automatically during rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
knitr::opts_knit$set(root.dir = chapter_dir)
```

> This chapter uses hidden prerequisite setup from an earlier chapter so the later code can run in order.

```{r prerequisites, include=FALSE}
# Prerequisite code adapted from Chapter 02.
    df <- read.csv(file.path("data", "apartment_price_data.csv"))

    xbar <- mean(df$living_area)
    ybar <- mean(df$price)

    is.na(df$price)

    !is.na(df$price) & !is.na(df$living_area)

    df <- df[!is.na(df$price) & !is.na(df$living_area), ]

    xbar <- mean(df$living_area)
    ybar <- mean(df$price)
    n <- nrow(df)

      s_xy <- 1 / (n - 1) * sum((df$living_area - xbar) *
                                (df$price - ybar))

      s_x2 <- 1 / (n - 1) * sum((df$living_area - xbar)^2)
      s_y2 <- 1 / (n - 1) * sum((df$price - ybar)^2)

      corr_xy <- s_xy / (sqrt(s_x2) * sqrt(s_y2))

      corr_xy <- cov(df$living_area, df$price) / 
                 (sd(df$living_area) * sd(df$price))

      corr_xy <- cor(df$living_area, df$price)

    df$Rx <- rank(df$living_area)
    df$Ry <- rank(df$price)

    mean_Rx <- (n + 1) / 2
    mean_Ry <- (n + 1) / 2

    covRxRy <- sum((df$Rx - mean_Rx)*(df$Ry-mean_Ry)) / (n - 1)
   sRx <- sqrt(sum((df$Rx - mean_Rx)^2) / (n - 1))
sRy <- sqrt(sum((df$Ry - mean_Ry)^2) / (n - 1))

    rS <- covRxRy / (sRx * sRy)

    rS_alt <- cor(df$Rx, df$Ry, method = "spearman")

  beta_1_hat <- s_xy / s_x2
  beta_0_hat <- ybar - xbar * beta_1_hat

  y_hat <- beta_0_hat + beta_1_hat * df$living_area

 u_hat <- df$price - y_hat

  RSS <- sum(u_hat^2)
  TSS <- sum((df$price - ybar)^2)
  R2 <- 1 - RSS / TSS

 ols_model <- lm(price ~ living_area, data=df)

 summary(ols_model)

```

## Inference for regression in R

### Box 01

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-01}
 s2_epsilon <- RSS / (n - 2)
 SSX <- sum((df$living_area - mean(df$living_area))^2)
 var_hat_beta_1_hat <- s2_epsilon / SSX
```

### Box 02

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-02}
 beta_1_H0 <- 0
 se_beta_1_hat <- sqrt(var_hat_beta_1_hat)
 t_stat <- (beta_1_hat - beta_1_H0) / se_beta_1_hat
 p_val <- 2 * (1 - pt(abs(t_stat), df=n-2))
```

### Box 03

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-03}
    alpha <- 0.1
    t_crit <- qt(1-alpha/2, df=n-2)
    lb <- beta_1_hat - t_crit * se_beta_1_hat
    ub <- beta_1_hat + t_crit * se_beta_1_hat
```

### Box 04

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-04}
 sum_ols_model <- summary(ols_model)
```

### Box 05

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-05}
 reg_table <- sum_ols_model$coefficients
```

### Box 06

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-06}
 reg_table[2, 3]
```

### Box 07

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-07}
 confint(ols_model, level=0.9)
```

### Box 08

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-08}
 ols_model2 <- lm(price ~ new_production, data=df)
```

### Box 09

Textbook context: Section: Inference to the population with spherical error terms.

```{r runnable-inference-for-regression-in-r-09}
    confint(ols_model2, level=0.9)
```

## Robust inference in R

### Box 10

Textbook context: Section: Asymptotic inference to the population.

```{r runnable-robust-inference-in-r-10}
 dx2 <- (df$living_area - mean(df$living_area))^2
```

### Box 11

Textbook context: Section: Asymptotic inference to the population.

```{r runnable-robust-inference-in-r-11}
 var_beta_1_var_r <- sum(dx2 * u_hat^2) / (sum(dx2))^2
```

### Box 12

Textbook context: Section: Asymptotic inference to the population.

```{r runnable-robust-inference-in-r-12}
 se_beta_1_hat_r <- sqrt(var_beta_1_var_r)
```

### Box 13

Textbook context: Section: Asymptotic inference to the population.

```{r runnable-robust-inference-in-r-13}
    library("lmtest")
    library("sandwich")
```

### Box 14

Textbook context: Section: Asymptotic inference to the population.

```{r runnable-robust-inference-in-r-14}
 coeftest(ols_model, vcov=vcovHC(ols_model, type="HC0"))
```
