---
title: "Chapter 03: Basic probability theory and statistical inference"
subtitle: "Runnable workbook for Regression with R and Python"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This runnable workbook supplements the reference workbook. It adds setup code and lightweight fixes so the chapter can usually be rendered and run end-to-end.

- Runnable companion script: [R/chapter03.R](R/chapter03.R)
- Reference workbook: [chapter03_workbook.Rmd](chapter03_workbook.Rmd)
- Textbook chapter: `Basic probability theory and statistical inference`
- Package installation lines are shown as comments and are not run automatically during rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
knitr::opts_knit$set(root.dir = chapter_dir)
```

## Monte Carlo simulation in R

### Box 01

Textbook context: Section: Estimators.

```{r runnable-monte-carlo-simulation-in-r-01}
    n <- 50
    dice_throws <- sample(1:6, n, replace=TRUE)
    ybar <- mean(dice_throws)
```

### Box 02

Textbook context: Section: Estimators.

```{r runnable-monte-carlo-simulation-in-r-02}
    nr_samples <- 10000
    n <- 50
    ybar_dist <- rep(NA, nr_samples)
    for (i in 1:nr_samples){
      dice_throws <- sample(1:6, n, replace=TRUE)
      ybar_dist[i] <- mean(dice_throws)
      print(i)
    }
```

### Box 03

Textbook context: Section: Estimators.

```{r runnable-monte-carlo-simulation-in-r-03}
 hist(ybar_dist, breaks=50)
```

## Hypothesis tests and confidence intervals with R

### Box 04

Textbook context: Section: Hypothesis testing and confidence intervals | Subsection: Confidence interval.

```{r runnable-hypothesis-tests-and-confidence-intervals-with-r-04}
    dice_throws <- c(rep(1, 13), rep(2, 7), rep(3, 8),
                     rep(4, 9), rep(5, 9), rep(6, 4))
```

### Box 05

Textbook context: Section: Hypothesis testing and confidence intervals | Subsection: Confidence interval.

```{r runnable-hypothesis-tests-and-confidence-intervals-with-r-05}
    n <- length(dice_throws)
    mean_dice_throws <- mean(dice_throws)
    var_dice_throws <- 
      sum((dice_throws - mean_dice_throws)^2) / (n - 1)
```

### Box 06

Textbook context: Section: Hypothesis testing and confidence intervals | Subsection: Confidence interval.

```{r runnable-hypothesis-tests-and-confidence-intervals-with-r-06}
    se_dice_throws <- sqrt(var_dice_throws / n)
    t_stat <- (mean_dice_throws - 3.5) / se_dice_throws
```

### Box 07

Textbook context: Section: Hypothesis testing and confidence intervals | Subsection: Confidence interval.

```{r runnable-hypothesis-tests-and-confidence-intervals-with-r-07}
    p_value <- 2 * (1 - pt(abs(t_stat), df=n-1))
```

### Box 08

Textbook context: Section: Hypothesis testing and confidence intervals | Subsection: Confidence interval.

```{r runnable-hypothesis-tests-and-confidence-intervals-with-r-08}
    alpha <- 0.05
    t_crit <- qt(1-alpha/2, n-1)
    ci_lower <- mean_dice_throws - t_crit * se_dice_throws
    ci_upper <- mean_dice_throws + t_crit * se_dice_throws
```

### Box 09

Textbook context: Section: Hypothesis testing and confidence intervals | Subsection: Confidence interval.

```{r runnable-hypothesis-tests-and-confidence-intervals-with-r-09}
 t.test(dice_throws, mu=3.5)
```
