---
title: "Chapter 10: Prediction"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Inference"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter10.R](R/chapter10.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Prediction`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Example with training and test data

### Box 01

Textbook context: Section: Training data, test data and cross-validation.

```{r example-with-training-and-test-data-01}
    set.seed(12)
    n <- nrow(df)
    train_share <- .8
    train_ind <- sample(1:n, train_share * n, replace=FALSE)
    test_ind <- setdiff(1:n, train_ind)
    df_train <- df[train_ind, ]
    df_test <- df[test_ind, ]
```

## Ridge and lasso in R

### Box 02

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r ridge-and-lasso-in-r-02}
 # install.packages("glmnet")
 library(glmnet)
```

### Box 03

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r ridge-and-lasso-in-r-03}
  df <- read.csv("apartment_price_data.csv")
  df <- df[complete.cases(df), ]
```

### Box 04

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r ridge-and-lasso-in-r-04}
 dfX <- df[, names(df) != "price"]
```

### Box 05

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r ridge-and-lasso-in-r-05}
 X <- makeX(dfX)
```

### Box 06

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r ridge-and-lasso-in-r-06}
 yhat <- predict(ridge_model, newx = X)
```

### Box 07

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r ridge-and-lasso-in-r-07}
 plot(ridge_model_cv)
```

## Regression tree in R

### Box 08

Textbook context: Section: Tree-based regression models.

```{r regression-tree-in-r-08}
  # install.packages("rpart")
  # install.packages("rpart.plot")
  library(rpart)
  library(rpart.plot)
```

### Box 09

Textbook context: Section: Tree-based regression models.

```{r regression-tree-in-r-09}
 df <- read.csv("apartment_price_data.csv")
```

### Box 10

Textbook context: Section: Tree-based regression models.

```{r regression-tree-in-r-10}
 tree_model <- rpart(price ~ living_area + monthly_fee,
 data=df, control=rpart.control(
 minsplit = 20, minbucket = 5, cp=0))
```

### Box 11

Textbook context: Section: Tree-based regression models.

```{r regression-tree-in-r-11}
  rpart.plot(tree_model)
```

### Box 12

Textbook context: Section: Tree-based regression models.

```{r regression-tree-in-r-12}
 pruned_tree <- prune(tree_model, cp=.03)
```

### Box 13

Textbook context: Section: Tree-based regression models.

```{r regression-tree-in-r-13}
 printcp(tree_model)
```

## Comparison of the different prediction models

### Box 14

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-14}
 year_list10 <- seq(from=1900, to=2010, by=10)
```

### Box 15

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-15}
  room_vars <- ls(df, pat="^room_size_|^city_area_room_size_")
```

### Box 16

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-16}
 build_decade_vars <- ls(df,
 pat="^build_decade_|^city_area_build_decade_")
 build_year_vars <- ls(df,
 pat="^build_year_|^city_area_build_year_")
```

### Box 17

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-17}
  rest_vars <- c("living_area", "new_production",
                 "monthly_fee", "city_area", "elevator",
                 "elevator_missing", "living_area2",
                 "monthly_fee2")
  tree_vars <- c("living_area", "new_production",
                 "monthly_fee", "city_area", "elevator",
                 "number_of_rooms", "build_year")
```

### Box 18

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-18}
 varlist_model1 <- c(room_vars, build_year_vars, rest_vars)
 varlist_model10 <- c(room_vars, build_decade_vars, rest_vars)
```

### Box 19

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-19}
 formOLS1 <- paste(varlist_model1, collapse="+")
 formOLS1 <- paste("price ~ ", formOLS1)
```

### Box 20

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-20}
  formOLS10 <- paste("price ~ ",
    paste(varlist_model10, collapse="+"))
  formCART <- paste("price ~ ",
    paste(tree_vars, collapse="+"))
```

### Box 21

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-21}
 set.seed(12)
 n <- nrow(df)
 ind_train <- sample(1:n, 4/5 * n, replace=FALSE)
 ind_test <- setdiff(1:n, ind_train)
```

### Box 22

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-22}
  ols_model1 <- lm(formOLS1, data=df[ind_train, ])
  ridge_model1 <- cv.glmnet(
    X1[ind_train, ], Y[ind_train], alpha=0)
  lasso_model1 <- cv.glmnet(
    X1[ind_train, ], Y[ind_train], alpha=1)

  ols_model10 <- lm(formOLS10, data=df[ind_train, ])
  ridge_model10 <- cv.glmnet(
    X10[ind_train, ], Y[ind_train], alpha=0)
  lasso_model10 <- cv.glmnet(
    X10[ind_train, ], Y[ind_train], alpha=1)

    CART_model = rpart(formCART, data=df[ind_train, ],
                     control=rpart.control(
                      minsplit = 20, minbucket = 5, cp=0))
  c <- printcp(CART_model)
  s <- c[, "xerror"]
  alpha <- c[, "CP"][s == min(s)]
  pruned_tree <- prune(CART_model, cp=alpha)
```

### Box 23

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-23}
  yhat_ols1 <- predict(ols_model1, newdata=df)
  yhat_ols10 <- predict(ols_model10, newdata=df)
  yhat_ridge1 <- predict(ridge_model1, newx=X1,
                         s="lambda.min")
  yhat_ridge10 <- predict(ridge_model10, newx=X10,
                          s="lambda.min")
  yhat_lasso1 <- predict(lasso_model1, newx=X1,
                         s="lambda.min")
  yhat_lasso10 <- predict(lasso_model10, newx=X10,
                          s="lambda.min")
  yhat_tree <- predict(pruned_tree, newdata = df)
```

### Box 24

Textbook context: Section: Example: housing prices.

```{r comparison-of-the-different-prediction-models-24}
 YHAT_models <- list(yhat_ols1, yhat_ridge1, yhat_lasso1,
 yhat_ols10, yhat_ridge10, yhat_lasso10,
 yhat_tree)
 MSE_train <- c()
 MSE_test <- c()
 for (yhat in YHAT_models) {
 mse_train <- mean((Y[ind_train] - yhat[ind_train])^2)
 mse_test <- mean((Y[ind_test] - yhat[ind_test])^2)
 MSE_train <- c(MSE_train, mse_train)
 MSE_test <- c(MSE_test, mse_test)
 }
```
