---
title: "Chapter 10: Prediction and Nonparametric regression"
subtitle: "Runnable workbook for Regression with R and Python"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This runnable workbook supplements the reference workbook. It adds setup code and lightweight fixes so the chapter can usually be rendered and run end-to-end.

- Runnable companion script: [R/chapter10.R](R/chapter10.R)
- Reference workbook: [chapter10_workbook.Rmd](chapter10_workbook.Rmd)
- Textbook chapter: `Prediction and Nonparametric regression`
- Package installation lines are shown as comments and are not run automatically during rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
knitr::opts_knit$set(root.dir = chapter_dir)
```

## Example with training and test data

### Box 01

Textbook context: Section: Training data, test data and cross-validation.

```{r runnable-example-with-training-and-test-data-01}
 df <- read.csv(file.path("data", "apartment_price_data.csv"))
 df$living_area2 <- df$living_area^2
 df$living_area3 <- df$living_area^3
 df$living_area4 <- df$living_area^4
 df$living_area5 <- df$living_area^5
 df$living_area6 <- df$living_area^6
 df$living_area7 <- df$living_area^7
 df$living_area8 <- df$living_area^2
 df$living_area9 <- df$living_area^9
 df$living_area10 <- df$living_area^10
```

### Box 02

Textbook context: Section: Training data, test data and cross-validation.

```{r runnable-example-with-training-and-test-data-02}
    for (i in 2:10) {
      df[, paste0("living_area", i)] <- df$living_area^i
    }
```

### Box 03

Textbook context: Section: Training data, test data and cross-validation.

```{r runnable-example-with-training-and-test-data-03}
    set.seed(12)
    n <- nrow(df)
    train_share <- .8
    train_ind <- sample(1:n, train_share * n, replace=FALSE)
    test_ind <- setdiff(1:n, train_ind)
    df_train <- df[train_ind, ]
    df_test <- df[test_ind, ]
```

### Box 04

Textbook context: Section: Training data, test data and cross-validation.

```{r runnable-example-with-training-and-test-data-04}
    ols_model <- lm(price ~ living_area, data=df_train)
    yhat_train <- predict(ols_model)
    yhat_test <- predict(ols_model, newdata=df_test)
    mse_train <- mean((df_train$price - yhat_train)^2)
    mse_test <- mean((df_test$price - yhat_test)^2)
```

### Box 05

Textbook context: Section: Training data, test data and cross-validation.

```{r runnable-example-with-training-and-test-data-05}
    f <- "price ~ living_area"
    mse_train_list <- c(mse_train)
    mse_test_list <- c(mse_test)
    for (i in 2:10) {
      f <- paste0(f, " + living_area", i)
      ols_model <- lm(as.formula(f), data=df_train)
      yhat_train <- predict(ols_model)
      yhat_test <- predict(ols_model, newdata=df_test)
      mse_train_list <- append(mse_train_list, 
        mean((df_train$price - yhat_train)^2))
      mse_test_list <- append(mse_test_list,
        mean((df_test$price - yhat_test)^2))
    }
```

## Cross-validation in R

### Box 06

Textbook context: Section: Training data, test data and cross-validation | Subsection: Cross-validation.

```{r runnable-cross-validation-in-r-06}
    df <- read.csv(file.path("data", "apartment_price_data.csv"))
    n <- nrow(df)
    m <- 5
    shuffle_ind <- sample(1:n, n, replace=FALSE)
    fold_indexes <- cut(shuffle_ind, breaks=m, labels=FALSE)
    MSE_hat <- rep(NA, m)
    for (i in 1:m){
      leave_out_ind <- (1:n)[fold_indexes == i]
      leave_in_ind <- setdiff(1:n, leave_out_ind)
      train_df <- df[leave_in_ind, ]
      test_df <- df[leave_out_ind, ]
      ols_train <- lm(price ~ living_area, data=train_df)
      pred <- predict(ols_train, test_df)
      MSE_hat[i] <- mean((test_df$price - pred)^2)
    }
    mean(MSE_hat)
```

## Ridge and lasso in R

### Box 07

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-07}
# install.packages("glmnet")  # install manually if needed
 library(glmnet)
```

### Box 08

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-08}
  df <- read.csv(file.path("data", "apartment_price_data.csv"))
  df <- df[complete.cases(df), ]
```

### Box 09

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-09}
 dfX <- df[, names(df) != "price"]
```

### Box 10

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-10}
 dfX$build_year <- as.factor(dfX$build_year)
 dfX$number_of_rooms <- as.factor(dfX$number_of_rooms)
```

### Box 11

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-11}
 X <- makeX(dfX)
```

### Box 12

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-12}
  ridge_model <- glmnet(X, df$price, alpha=0)
```

### Box 13

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-13}
 ridge_model <- glmnet(X, df$price, alpha=0, lambda=.1)
```

### Box 14

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-14}
 yhat <- predict(ridge_model, newx = X)
```

### Box 15

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-15}
  ridge_model_cv <- cv.glmnet(X, df$price, alpha=0)
```

### Box 16

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-16}
 plot(ridge_model_cv)
```

### Box 17

Textbook context: Section: Predictions with many independent variables | Subsection: Scale dependence with ridge and lasso regression.

```{r runnable-ridge-and-lasso-in-r-17}
 lasso_model_cv <- cv.glmnet(X, df$price, alpha=1)
```

## Regression tree in R

### Box 18

Textbook context: Section: Tree-based regression models.

```{r runnable-regression-tree-in-r-18}
# install.packages("rpart")  # install manually if needed
# install.packages("rpart.plot")  # install manually if needed
  library(rpart)
  library(rpart.plot)
```

### Box 19

Textbook context: Section: Tree-based regression models.

```{r runnable-regression-tree-in-r-19}
 df <- read.csv(file.path("data", "apartment_price_data.csv"))
```

### Box 20

Textbook context: Section: Tree-based regression models.

```{r runnable-regression-tree-in-r-20}
 tree_model <- rpart(price ~ living_area + monthly_fee,
 data=df, control=rpart.control(
 minsplit = 20, minbucket = 5, cp=0))
```

### Box 21

Textbook context: Section: Tree-based regression models.

```{r runnable-regression-tree-in-r-21}
  rpart.plot(tree_model)
```

### Box 22

Textbook context: Section: Tree-based regression models.

```{r runnable-regression-tree-in-r-22}
 pruned_tree <- prune(tree_model, cp=.03)
```

### Box 23

Textbook context: Section: Tree-based regression models.

```{r runnable-regression-tree-in-r-23}
 printcp(tree_model)
```

## Comparison of the different prediction models

### Box 24

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-24}
 library(glmnet)
 library(rpart)
 df <- read.csv(file.path("data", "apartment_price_data.csv"))
 df$elevator_missing <- ifelse(is.na(df$elevator), 1, 0)
 df$elevator[is.na(df$elevator)] <- 0
 df$living_area2 <- df$living_area^2
 df$monthly_fee2 <- df$monthly_fee^2
 df <- df[complete.cases(df), ]
```

### Box 25

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-25}
  for (room_size in 2:6){
    df[, paste0("room_size_", room_size)] <- ifelse(
      df$number_of_rooms == room_size, 1, 0)
    df[, paste0("city_area_room_size_", room_size)] <- (
      df$city_area * 
      df[, paste0("room_size_", room_size)]) 
  }
```

### Box 26

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-26}
 year_list10 <- seq(from=1900, to=2010, by=10)
```

### Box 27

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-27}
  for (year in year_list10){
    df[, paste0("build_decade_", year)] <- ifelse(
      df$build_year >= year & df$build_year <= year + 9, 1, 0)
    df[, paste0("city_area_build_decade_", year)] <- (
      df$city_area * 
        df[, paste0("build_decade_", year)]) 
  }
```

### Box 28

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-28}
 year_list1 <- unique(df$build_year)
 for (year in year_list1){
 df[, paste0("build_year_", year)] <- ifelse(
 df$build_year == year, 1, 0)
 df[, paste0("city_area_build_year_", year)] <- (
 df$city_area *
 df[, paste0("build_year_", year)])
 }
```

### Box 29

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-29}
  room_vars <- grep("^room_size_|^city_area_room_size_", names(df), value=TRUE)
```

### Box 30

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-30}
build_decade_vars <- grep("^build_decade_|^city_area_build_decade_", names(df), value=TRUE)
build_year_vars <- grep("^build_year_|^city_area_build_year_", names(df), value=TRUE)
```

### Box 31

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-31}
  rest_vars <- c("living_area", "new_production",
                 "monthly_fee", "city_area", "elevator",
                 "elevator_missing", "living_area2",
                 "monthly_fee2")
  tree_vars <- c("living_area", "new_production",
                 "monthly_fee", "city_area", "elevator",
                 "number_of_rooms", "build_year")
```

### Box 32

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-32}
 varlist_model1 <- c(room_vars, build_year_vars, rest_vars)
 varlist_model10 <- c(room_vars, build_decade_vars, rest_vars)
```

### Box 33

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-33}
 X1 <- as.matrix(df[, varlist_model1])
 X10 <- as.matrix(df[, varlist_model10])
 Y <- df$price
```

### Box 34

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-34}
 formOLS1 <- paste(varlist_model1, collapse="+")
 formOLS1 <- paste("price ~ ", formOLS1)
```

### Box 35

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-35}
  formOLS10 <- paste("price ~ ",
    paste(varlist_model10, collapse="+"))
  formCART <- paste("price ~ ",
    paste(tree_vars, collapse="+"))
```

### Box 36

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-36}
 set.seed(12)
 n <- nrow(df)
 ind_train <- sample(1:n, 4/5 * n, replace=FALSE)
 ind_test <- setdiff(1:n, ind_train)
```

### Box 37

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-37}
  ols_model1 <- lm(as.formula(formOLS1), data=df[ind_train, ])
  ridge_model1 <- cv.glmnet(
    X1[ind_train, ], Y[ind_train], alpha=0)
  lasso_model1 <- cv.glmnet(
    X1[ind_train, ], Y[ind_train], alpha=1)

  ols_model10 <- lm(as.formula(formOLS10), data=df[ind_train, ])
  ridge_model10 <- cv.glmnet(
    X10[ind_train, ], Y[ind_train], alpha=0)
  lasso_model10 <- cv.glmnet(
    X10[ind_train, ], Y[ind_train], alpha=1)

    CART_model = rpart(formCART, data=df[ind_train, ], 
                     control=rpart.control(
                      minsplit = 20, minbucket = 5, cp=0))
  c <- printcp(CART_model)
  s <- c[, "xerror"]
  alpha <- c[which.min(s), "CP"]
  pruned_tree <- prune(CART_model, cp=alpha)
```

### Box 38

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-38}
  yhat_ols1 <- predict(ols_model1, newdata=df)
  yhat_ols10 <- predict(ols_model10, newdata=df)
  yhat_ridge1 <- predict(ridge_model1, newx=X1,
                         s="lambda.min")
  yhat_ridge10 <- predict(ridge_model10, newx=X10,
                          s="lambda.min")
  yhat_lasso1 <- predict(lasso_model1, newx=X1,
                         s="lambda.min")
  yhat_lasso10 <- predict(lasso_model10, newx=X10,
                          s="lambda.min")
  yhat_tree <- predict(pruned_tree, newdata = df)
```

### Box 39

Textbook context: Section: Example: housing prices.

```{r runnable-comparison-of-the-different-prediction-models-39}
 YHAT_models <- list(yhat_ols1, yhat_ridge1, yhat_lasso1,
 yhat_ols10, yhat_ridge10, yhat_lasso10,
 yhat_tree)
 MSE_train <- c()
 MSE_test <- c()
 for (yhat in YHAT_models) {
 mse_train <- mean((Y[ind_train] - yhat[ind_train])^2)
 mse_test <- mean((Y[ind_test] - yhat[ind_test])^2)
 MSE_train <- c(MSE_train, mse_train)
 MSE_test <- c(MSE_test, mse_test)
 }
```
