---
title: "Chapter 04: Correlation and inference about a population"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter04.R](R/chapter04.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Correlation and inference about a population`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Calculating confidence intervals for correlation in R

### Box 01

Textbook context: Chapter context unavailable.

```{r calculating-confidence-intervals-for-correlation-in-r-01}
  df <- read.csv("apartment_price_data.csv")
  corrXY <- cor(df$living_area, df$price)
  Z <- 1/2 * log((1+corrXY)/(1-corrXY))
  alpha <- 0.05
  z_alpha_2 <- qnorm(1-alpha/2)
  Z_L <- Z - z_alpha_2 * sqrt(1/(1000-3))
  Z_U <- Z + z_alpha_2 * sqrt(1/(1000-3))
  CI_L <- (exp(2*Z_L) - 1)/(exp(2*Z_L) + 1)
  CI_U <- (exp(2*Z_U) - 1)/(exp(2*Z_U) + 1)
```

### Box 02

Textbook context: Chapter context unavailable.

```{r calculating-confidence-intervals-for-correlation-in-r-02}
  cor.test(df$living_area, df$price, conf.level=0.95)
```
