---
title: "Chapter 04: Correlation and inference about a population"
subtitle: "Runnable workbook for Regression with R and Python"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This runnable workbook supplements the reference workbook. It adds setup code and lightweight fixes so the chapter can usually be rendered and run end-to-end.

- Runnable companion script: [R/chapter04.R](R/chapter04.R)
- Reference workbook: [chapter04_workbook.Rmd](chapter04_workbook.Rmd)
- Textbook chapter: `Correlation and inference about a population`
- Package installation lines are shown as comments and are not run automatically during rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
knitr::opts_knit$set(root.dir = chapter_dir)
```

## Calculating confidence intervals for correlation in R

### Box 01

Textbook context: Chapter context unavailable.

```{r runnable-calculating-confidence-intervals-for-correlation-in-r-01}
  df <- read.csv(file.path("data", "apartment_price_data.csv"))
  corrXY <- cor(df$living_area, df$price)
  Z <- 1/2 * log((1+corrXY)/(1-corrXY))
  alpha <- 0.05
  z_alpha_2 <- qnorm(1-alpha/2)
  Z_L <- Z - z_alpha_2 * sqrt(1/(1000-3))
  Z_U <- Z + z_alpha_2 * sqrt(1/(1000-3))
  CI_L <- (exp(2*Z_L) - 1)/(exp(2*Z_L) + 1)
  CI_U <- (exp(2*Z_U) - 1)/(exp(2*Z_U) + 1)
```

### Box 02

Textbook context: Chapter context unavailable.

```{r runnable-calculating-confidence-intervals-for-correlation-in-r-02}
  cor.test(df$living_area, df$price, conf.level=0.95)
```
