---
title: "Chapter 12: Causal analyses"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter12.R](R/chapter12.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Causal analyses`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## RD approach in R

### Box 01

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r rd-approach-in-r-01}
install.packages(c("sandwich", "lmtest", "ivreg", "rdrobust"))
```

### Box 02

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r rd-approach-in-r-02}
 df <- read.csv("municip_data.csv")
 df <- df[df$municip_name != "Gotland", ]
 df$rv <- df$share_seats_left_last_election - .5
 df$change_tax_rate <- df$tax_rate - df$tax_rate_4_years_back
 df <- df[!is.na(df$change_tax_rate), ]
 df$Z <- ifelse(df$rv >= 0, 1, 0)
 df$Zrv <- df$Z * df$rv
```

### Box 03

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r rd-approach-in-r-03}
  h <- .05
  rd_fs <- lm(left_coalition_last_term ~ Z + rv + Zrv,
              data=df[abs(df$rv) <= h, ])
  rd_rf <- lm(change_tax_rate ~ Z + rv + Zrv,
              data=df[abs(df$rv) <= h, ])
  Dhat <- predict(rd_fs)
  rd_2sls <- lm(change_tax_rate ~ Dhat + rv + Zrv,
                data=df[abs(df$rv) <= h, ])
  library(sandwich)
  library(lmtest)
  coeftest(rd_fs, vcov=vcovHC(rd_fs, type="HC0"))
  coeftest(rd_rf, vcov=vcovHC(rd_rf, type="HC0"))
  summary(rd_2sls)
```

### Box 04

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r rd-approach-in-r-04}
  library(ivreg)
  rd_2sls_correct <- ivreg(change_tax_rate ~ 
      left_coalition_last_term + rv + Zrv | Z + rv + Zrv,
      data=df[abs(df$rv) <= h, ])
  coeftest(rd_2sls_correct, vcov=vcovHC(rd_2sls_correct,
                                        type="HC0"))
```

### Box 05

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r rd-approach-in-r-05}
 install.packages("rdrobust")
 library(rdrobust)
 rd_fs_alt <- rdrobust(df$left_coalition_last_term,
 df$share_seats_left_last_election,
 c=.5, kernel="uniform", h=h)
```

### Box 06

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r rd-approach-in-r-06}
  rd_2sls_alt <- rdrobust(df$change_tax_rate,
    df$share_seats_left_last_election,
    c=.5, kernel="uniform", h=h,
    fuzzy=df$left_coalition_last_term)
```
