---
title: "Chapter 12: Causal analyses"
subtitle: "Runnable workbook for Regression with R and Python"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This runnable workbook supplements the reference workbook. It adds setup code and lightweight fixes so the chapter can usually be rendered and run end-to-end.

- Runnable companion script: [R/chapter12.R](R/chapter12.R)
- Reference workbook: [chapter12_workbook.Rmd](chapter12_workbook.Rmd)
- Textbook chapter: `Causal analyses`
- Package installation lines are shown as comments and are not run automatically during rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
knitr::opts_knit$set(root.dir = chapter_dir)
```

## RD approach in R

### Box 01

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r runnable-rd-approach-in-r-01}
# install.packages(c("sandwich", "lmtest", "ivreg", "rdrobust"))  # install manually if needed
```

### Box 02

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r runnable-rd-approach-in-r-02}
 df <- read.csv(file.path("data", "municip_data.csv"))
 df <- df[df$municip_name != "Gotland", ]
 df$rv <- df$share_seats_left_last_election - .5
 df$change_tax_rate <- df$tax_rate - df$tax_rate_4_years_back
 df <- df[!is.na(df$change_tax_rate), ]
 df$Z <- ifelse(df$rv >= 0, 1, 0)
 df$Zrv <- df$Z * df$rv
```

### Box 03

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r runnable-rd-approach-in-r-03}
  h <- .05
  rd_fs <- lm(left_coalition_last_term ~ Z + rv + Zrv,
              data=df[abs(df$rv) <= h, ])
  rd_rf <- lm(change_tax_rate ~ Z + rv + Zrv,
              data=df[abs(df$rv) <= h, ])
  Dhat <- predict(rd_fs)
  rd_2sls <- lm(change_tax_rate ~ Dhat + rv + Zrv,
                data=df[abs(df$rv) <= h, ])
  library(sandwich)
  library(lmtest)
  coeftest(rd_fs, vcov=vcovHC(rd_fs, type="HC0"))
  coeftest(rd_rf, vcov=vcovHC(rd_rf, type="HC0"))
  summary(rd_2sls)
```

### Box 04

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r runnable-rd-approach-in-r-04}
  library(ivreg)
  rd_2sls_correct <- ivreg(change_tax_rate ~ 
      left_coalition_last_term + rv + Zrv | Z + rv + Zrv,
      data=df[abs(df$rv) <= h, ])
  coeftest(rd_2sls_correct, vcov=vcovHC(rd_2sls_correct,
                                        type="HC0"))
```

### Box 05

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r runnable-rd-approach-in-r-05}
# install.packages("rdrobust")  # install manually if needed
 library(rdrobust)
 rd_fs_alt <- rdrobust(df$left_coalition_last_term,
 df$share_seats_left_last_election,
 c=.5, kernel="uniform", h=h)
```

### Box 06

Textbook context: Section: The regression discontinuity approach | Subsection: Example in R with municipal data.

```{r runnable-rd-approach-in-r-06}
  rd_2sls_alt <- rdrobust(df$change_tax_rate,
    df$share_seats_left_last_election,
    c=.5, kernel="uniform", h=h,
    fuzzy=df$left_coalition_last_term)
```
