---
title: "Chapter 13: Nonparametric Regression"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter13.R](R/chapter13.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Nonparametric Regression`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Nonparametric RD in R (local linear + triangular kernel)

### Box 01

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-01}
install.packages(c("sandwich", "lmtest", "ivreg", "rdrobust"))
```

### Box 02

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-02}
df <- read.csv("municip_data.csv")
df <- df[df$municip_name != "Gotland", ]

c0 <- .5
df$rv <- df$share_seats_left_last_election - c0
df$change_tax_rate <- df$tax_rate - df$tax_rate_4_years_back
df <- df[!is.na(df$change_tax_rate), ]

df$Z   <- ifelse(df$rv >= 0, 1, 0)   # cutoff indicator
df$Zrv <- df$Z * df$rv              # slope interaction
```

### Box 03

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-03}
tri <- function(u) pmax(0, 1 - abs(u))

h <- .10
df_h <- df[abs(df$rv) <= h, ]
df_h$w <- tri(df_h$rv / h)
```

### Box 04

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-04}
rd_fs_np <- lm(left_coalition_last_term ~ Z + rv + Zrv,
               data = df_h, weights = w)

rd_rf_np <- lm(change_tax_rate ~ Z + rv + Zrv,
               data = df_h, weights = w)

library(sandwich)
library(lmtest)

coeftest(rd_fs_np, vcov = vcovHC(rd_fs_np, type = "HC0"))
coeftest(rd_rf_np, vcov = vcovHC(rd_rf_np, type = "HC0"))
```

### Box 05

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-05}
delta_hat <- coef(rd_fs_np)["Z"]
alpha_hat <- coef(rd_rf_np)["Z"]
wald_hat  <- alpha_hat / delta_hat

library(ivreg)

rd_fuzzy_np <- ivreg(change_tax_rate ~ left_coalition_last_term + rv + Zrv |
                       Z + rv + Zrv,
                     data = df_h, weights = w)

coeftest(rd_fuzzy_np, vcov = vcovHC(rd_fuzzy_np, type = "HC0"))

c(wald_hat = wald_hat,
  iv_hat   = coef(rd_fuzzy_np)["left_coalition_last_term"])
```

### Box 06

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-06}
fig_dir <- "pdf"
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

binwidth <- .01
df$bin <- pmin(floor(df$share_seats_left_last_election / binwidth) * binwidth,
               1 - binwidth) + binwidth/2

agg_fs <- aggregate(left_coalition_last_term ~ bin, data = df, mean)
agg_rf <- aggregate(change_tax_rate ~ bin, data = df, mean)

xlim <- c(c0 - h, c0 + h)

## helper: fitted values on each side from the interacted local-linear model
fitted_side <- function(b, x, side = c("left","right")) {
  side <- match.arg(side)
  rv <- x - c0
  b0 <- b["(Intercept)"]; b1 <- b["Z"]; b2 <- b["rv"]; b3 <- b["Zrv"]
  if (side == "left")  return(b0 + b2 * rv)
  return((b0 + b1) + (b2 + b3) * rv)
}

## --- First stage plot ---
b_fs <- coef(rd_fs_np)
xL <- seq(xlim[1], c0, length.out = 200)
xR <- seq(c0, xlim[2], length.out = 200)

pdf(file.path(fig_dir, "rd_municip_fs_agg_np.pdf"),
    width = 4.2, height = 3.4, family = "serif")
par(mar = c(4.6, 4.8, 0.6, 0.6))

agg_fs2 <- agg_fs[agg_fs$bin >= xlim[1] & agg_fs$bin <= xlim[2], ]
plot(agg_fs2$bin, agg_fs2$left_coalition_last_term,
     pch = 16, cex = .85, col = "#F47939",
     xlab = "# seats left-wing parties",
     ylab = "Left-wing ruling coalition",
     xlim = xlim, ylim = c(0, 1), bty = "l")
abline(v = c0, lty = 2, lwd = 1.2)
lines(xL, fitted_side(b_fs, xL, "left"),  lwd = 2.2)
lines(xR, fitted_side(b_fs, xR, "right"), lwd = 2.2)
dev.off()

## --- Reduced form plot ---
b_rf <- coef(rd_rf_np)

pdf(file.path(fig_dir, "rd_municip_rf_agg_d_np.pdf"),
    width = 4.2, height = 3.4, family = "serif")
par(mar = c(4.6, 4.8, 0.6, 0.6))

agg_rf2 <- agg_rf[agg_rf$bin >= xlim[1] & agg_rf$bin <= xlim[2], ]
plot(agg_rf2$bin, agg_rf2$change_tax_rate,
     pch = 16, cex = .85, col = "#F47939",
     xlab = "# seats left-wing parties",
     ylab = "Change in tax rate",
     xlim = xlim, bty = "l")
abline(v = c0, lty = 2, lwd = 1.2)
lines(xL, fitted_side(b_rf, xL, "left"),  lwd = 2.2)
lines(xR, fitted_side(b_rf, xR, "right"), lwd = 2.2)
dev.off()
```

### Box 07

Textbook context: Section: Regression discontinuity | Subsection: Manipulation check: density around the cutoff.

```{r nonparametric-rd-in-r-local-linear-triangular-kernel-07}
library(rdrobust)

out_np <- rdrobust(y = df$change_tax_rate,
                   x = df$share_seats_left_last_election,
                   c = .5, kernel = "triangular",
                   fuzzy = df$left_coalition_last_term)
summary(out_np)
```
