---
title: "Chapter 09: Binary dependent variable"
subtitle: "Regression with R and Python: Description, Prediction, and Causal Analysis for Applied Social Science and Econometrics"
output:
  html_document:
    toc: true
    toc_depth: 2
---

This R Markdown workbook follows the chapter's code examples in textbook order. It is intended as a practical companion for advanced undergraduate and beginning master's students in applied social sciences and econometrics.

> This workbook is a reference companion: code chunks are shown in textbook order but are not executed during rendering.

- Companion R script: [R/chapter09.R](R/chapter09.R)
- Chapter overview: [README.md](README.md)
- Textbook chapter: `Binary dependent variable`

> Some textbook chunks install packages. If those packages are already installed on your machine, you can skip the installation lines when working through the notebook.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
chapter_dir <- dirname(normalizePath(knitr::current_input()))
data_dir <- file.path(chapter_dir, "data")
if (dir.exists(data_dir)) {
  knitr::opts_knit$set(root.dir = data_dir)
} else {
  knitr::opts_knit$set(root.dir = chapter_dir)
}
```

## Logistic regression in R

### Box 01

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-01}
 df <- read.csv("municip_data.csv")
 df <- df[df$year == 2022, ]
```

### Box 02

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-02}
    highest_quartile_limit <- quantile(df$tax_base, .75)
```

### Box 03

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-03}
 df$high_income <- ifelse(
 df$tax_base > highest_quartile_limit, 1, 0)
```

### Box 04

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-04}
    logit_model <- glm(df$high_income ~ share_tertiary_school,
                       family = binomial(link = "logit"),
                       data=df)
```

### Box 05

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-05}
 summary(logit_model)
```

### Box 06

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-06}
    df$lnpop <- log(df$pop)
    logit_model2 <- glm(df$high_income ~ 
      share_tertiary_school + lnpop,
      family = binomial(link = "logit"), data=df)
```

### Box 07

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-07}
    beta1hat <- logit_model2$coefficients[2]
    beta2hat <- logit_model2$coefficients[3]
```

### Box 08

Textbook context: Section: Estimation and inference under logistic regression model | Subsection: Assumptions about the error term.

```{r logistic-regression-in-r-08}
 exp(beta1hat * .01)
```

## Example in R with the probability of heart disease

### Box 09

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-09}
    df <- read.csv("heart_disease.csv")
    logit_model <- glm(df$heart_disease ~ blood_pressure,
                       family = binomial(link = "logit"),
                       data=df)
    summary(logit_model)
```

### Box 10

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-10}
    sum_logit_model <- summary(logit_model)$coefficients
    beta_1_hat <- sum_logit_model["blood_pressure", "Estimate"]
    se_beta_1_hat <- sum_logit_model[
      "blood_pressure", "Std. Error"]
    std_dev_blood_pressure <- sd(df$blood_pressure)
    alpha <- .05
    zstat <- qnorm(1-alpha/2)
    lb_beta_1_hat <- beta_1_hat - zstat * se_beta_1_hat
    ub_beta_1_hat <- beta_1_hat + zstat * se_beta_1_hat
    lb_odds_ratio <- exp(lb_beta_1_hat * std_dev_blood_pressure)
    ub_odds_ratio <- exp(ub_beta_1_hat * std_dev_blood_pressure)
    print(c(lb_odds_ratio, ub_odds_ratio))
```

### Box 11

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-11}
    logit_model2 <- glm(df$heart_disease ~ blood_pressure +
      male + age, family = binomial(link = "logit"), data=df)
    summary(logit_model2)
```

### Box 12

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-12}
    sum_logit_model2 <- summary(logit_model2)$coefficients
    beta_1_hat_b <- sum_logit_model2[
      "blood_pressure", "Estimate"]
    beta_2_hat_b <- sum_logit_model2["male", "Estimate"]
    beta_3_hat_b <- sum_logit_model2["age", "Estimate"]
    std_dev_age <- sd(df$age)
    odds_ratio_blood_pressure <- exp(beta_1_hat_b *
                                     std_dev_blood_pressure)
    odds_ratio_male <- exp(beta_2_hat_b * 1)
    odds_ratio_age <- exp(beta_3_hat_b * std_dev_age)
```

### Box 13

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-13}
    beta_hat_b <- sum_logit_model2[, "Estimate"][2:4]
    increases <- c(std_dev_blood_pressure, 1, std_dev_age)
    odds_ratios <- exp(beta_hat_b * increases)
```

### Box 14

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-14}
    beta_0_hat_b <- sum_logit_model2["(Intercept)", "Estimate"]
    pred_heart_disease <- 1 / (1 + exp(-(
      beta_0_hat_b + beta_1_hat_b * 140 + beta_2_hat_b * 1 +
      beta_3_hat_b * 50)))
```

### Box 15

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-15}
 pred_heart_disease_alt <- predict(logit_model2,
 data.frame(blood_pressure=140, male=1, age=50),
 type="response")
```

### Box 16

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-16}
    df$pred <- predict(logit_model2, type="response")
```

### Box 17

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-17}
    df_incr_blood_pressure <- df
    df_incr_blood_pressure$blood_pressure <- 
      df_incr_blood_pressure$blood_pressure + 
      std_dev_blood_pressure
    df_incr_blood_pressure$pred <- predict(logit_model2,
      newdata=df_incr_blood_pressure, type="response")
```

### Box 18

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-18}
 df$risk_ratio_blood_pressure <-
 df_incr_blood_pressure$pred / df$pred
 df$risk_diff_blood_pressure <-
 df_incr_blood_pressure$pred - df$pred
```

### Box 19

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-19}
 GRR_blood_pressure <- mean(df$risk_ratio_blood_pressure)
 GRD_blood_pressure <- mean(df$risk_diff_blood_pressure)
```

### Box 20

Textbook context: Section: Example: the risk of having a heart disease.

```{r example-in-r-with-the-probability-of-heart-disease-20}
    df_female <- df
    df_female$male <- 0
    df_male <- df
    df_male$male <- 1
    df_female$pred <- predict(logit_model2, newdata=df_female,
                              type="response")
    df_male$pred <- predict(logit_model2, newdata=df_male,
                            type="response")
    df$risk_ratio_male <- df_male$pred / df_female$pred
    df$risk_diff_male <- df_male$pred - df_female$pred
    GRR_male <- mean(df$risk_ratio_male)
    GRD_male <- mean(df$risk_diff_male)
```
